\documentclass[aspectratio=169]{beamer}
\usetheme{metropolis}
\usepackage{appendixnumberbeamer}
\usepackage{booktabs, hyperref}

\input{mgmt675-style}

\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows.meta, positioning, calc}

% Title info
\subtitle{MGMT 675: Generative AI for Finance}
\title{Module 1: AI in Finance}
\author{Kerry Back}
\date{}

\begin{document}

\maketitle

% ========================================
% SECTION 1: WHY AI MATTERS FOR FINANCE
% ========================================
\section{Why AI Matters for Finance}

\begin{frame}{From Makers to Checkers}
\centerline{Derek Waldron, JP Morgan Chief Analytics Officer}

What we're working towards is that every employee will have their own personalized AI assistant; every process is powered by AI agents, and every client experience has an AI concierge.

You'll still have people at the top who are managing and have relationships with clients, but many, many of the processes underneath are now being done by AI systems.

Workers would shift from being creators of reports or software updates, or `makers' \ldots\ to `checkers' or managers of AI agents doing that work.
\end{frame}

\begin{frame}{CFOs Are Going All In on AI}
\centerline{Deloitte Q4 2025 CFO Signals Survey --- 200 CFOs at \$1B+ companies}
\vspace{0.3cm}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{itemize}
  \item \textbf{87\%} say AI will be extremely or very important to finance operations in 2026
  \item \textbf{54\%} prioritize integrating AI agents in their finance departments
  \item \textbf{50\%} cite digital transformation of finance as their \#1 priority
\end{itemize}
\end{column}
\begin{column}{0.5\textwidth}
\begin{itemize}
  \item \textbf{49\%} prioritize automating processes to free employees for higher-value work
  \item Only \textbf{2\%} say AI won't be important
  \item CFO confidence at highest level since late 2021
\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}[shrink=5]{Case Study: HPE's ``Alfred''}
\centerline{Marie Myers, CFO of Hewlett Packard Enterprise (\#143 on Fortune 500) --- \href{https://fortune.com/2026/02/12/hpe-cfo-used-ai-transform-100-slide-monday-meeting/}{Fortune}, February 12, 2026}
\vspace{0.2cm}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{shadedbox}[title=\textbf{The Problem}]
\begin{itemize}\small
  \item Weekly 90-minute operational review required 100+ slides
  \item Hundreds of hours of preparation across business units
  \item No time left for forward-looking analysis
\end{itemize}
\end{shadedbox}
\end{column}
\begin{column}{0.5\textwidth}
\begin{shadedbox}[title=\textbf{The AI Solution}]
\begin{itemize}\small
  \item Built ``Alfred''---AI agents that pull, reconcile, and analyze data automatically
  \item \textbf{90\%} of manual prep eliminated; cycle time reduced \textbf{40\%}, costs down \textbf{25\%}
  \item 3,000+ finance employees being reskilled to build their own agents
\end{itemize}
\end{shadedbox}
\end{column}
\end{columns}
\vspace{0.2cm}
\centering
``The goal is for finance professionals to become \alert{masters of their own destiny} rather than casualties of automation.'' --- Marie Myers
\end{frame}

\begin{frame}{Something Big is Happening}
\centerline{Matt Shumer, CEO of HyperWrite AI, Feb.\ 2025}

The person who walks into a meeting and says ``I used AI to do this analysis in an hour instead of three days'' is going to be the most valuable person in the room. Not eventually. Right now.

\bigskip

Here's a simple commitment that will put you ahead of almost everyone: spend one hour a day experimenting with AI \ldots\ try to get it to do something new, something you're not sure it can handle. If you do this for the next six months, you will understand what's coming better than 99\% of the people around you.
\end{frame}

% ========================================
% SECTION 2: CLAUDE.AI AND ARTIFACTS
% ========================================
\section{Exploring Claude.ai}

\begin{frame}{Getting Started: Claude.ai}

Go to \texttt{claude.ai} in any browser. No installation required.  Icon at top left opens sidebar.

\textbf{What it can do:}
It's a chatbot with a code execution tool.  It can answer questions, explain concepts, brainstorm, draft text, summarize documents you paste in.  It can also generate charts, Excel files, Word docs, and PowerPoint decks.  Code runs in a \textbf{sandbox}---can \texttt{pip install} packages but has \alert{no internet access}.

\textbf{Getting started:}
\begin{enumerate}\small
  \item Go to \texttt{claude.ai} and sign in (or create an account)
  \item Type a question or upload a file
  \item Iterate: refine, ask follow-ups, request changes
\end{enumerate}
\end{frame}

\begin{frame}{Artifacts: Interactive Charts and Apps}

\begin{itemize}
  \item Claude can produce \textbf{interactive charts and apps} that appear in a side panel
  \item Choose Artifacts $\rightarrow$ Create an Artifact $\rightarrow$ apps or websites
  \item Click \textbf{Publish} to get a shareable link --- anyone can view and interact, no Claude account needed
\end{itemize}

\vspace{0.3cm}

\textbf{What Artifacts Can Build}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{itemize}\small
  \item Interactive charts with hover/zoom
  \item Financial calculators
  \item Scenario analysis with sliders
\end{itemize}
\end{column}
\begin{column}{0.5\textwidth}
\begin{itemize}\small
  \item ROI tools and comparisons
  \item Dashboards with filters
  \item Surveys and data collection
\end{itemize}
\end{column}
\end{columns}

\vspace{0.2cm}
\small\textbf{Limitation:} Artifacts are client-side only---no database, no persistence.  They are \alert{disposable by design}: generate a new one when needs change, rather than maintaining an old one.
\end{frame}

\begin{frame}{Artifact Example: Black-Scholes Calculator}
\textit{``Create a Black-Scholes option calculator with inputs for stock price, strike, rate, volatility, and expiration, and plot call and put prices as the stock price varies from 50\% to 150\% of the strike.''}

\vspace{0.3cm}
\textbf{Publish:} Publish the artifact and share the link.
\end{frame}

\begin{frame}{Artifacts Across Platforms}
All three major platforms let you describe what you want in plain English and get an interactive tool.

\begin{itemize}
  \item \textbf{Claude Artifacts} build React apps you can publish with one click---no account needed to view
  \item \textbf{ChatGPT Canvas} produces similar React apps with inline editing and shareable links
  \item \textbf{Gemini Canvas} generates HTML/CSS/JS and exports or copies the result
\end{itemize}

\vspace{0.3cm}
We use Claude, but the concepts transfer to any platform.
\end{frame}

% ========================================
% SECTION 3: EXPLORING CLAUDE FOR EXCEL
% ========================================
\section{Exploring Claude for Excel}

\begin{frame}{Claude for Excel Add-in}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{shadedbox}[title=\textbf{What It Does}]
\begin{itemize}\small
  \item AI sidebar that reads and edits your workbook
  \item Preserves formula dependencies
  \item Works with local files
\end{itemize}
\end{shadedbox}
\end{column}
\begin{column}{0.5\textwidth}
\begin{shadedbox}[title=\textbf{Capabilities}]
\begin{itemize}\small
  \item Ask questions about workbook content
  \item Build financial models from scratch
  \item Trace and fix formula errors
\end{itemize}
\end{shadedbox}
\end{column}
\end{columns}
\vspace{0.3cm}
Requires Claude Pro (\$20/mo) or a Claude Team/Enterprise plan.
\end{frame}

\begin{frame}{Installing the Add-in}
\begin{enumerate}
  \item Open \textbf{Excel} (desktop version, not the web app)
  \item Go to \textbf{Insert} $\rightarrow$ \textbf{Get Add-ins}
  \item Search for \textbf{``Claude by Anthropic''} and click \textbf{Add}
  \item Sign in with your Claude account when prompted
  \item The Claude sidebar appears on the right --- click the Claude icon in the \textbf{Home} ribbon tab to toggle it
\end{enumerate}
\vspace{0.3cm}
The add-in can see your entire workbook: sheet names, cell values, and formulas.  It edits cells directly --- you can undo any change with \texttt{Ctrl+Z}.
\end{frame}

\begin{frame}{Example: Loan Amortization Table}
Open a blank workbook with the Claude sidebar visible.  Type:
\vspace{0.2cm}
\begin{shadedbox}
\small\textit{``Create a loan amortization table for a \$300{,}000 mortgage at 6.5\% annual interest with a 30-year term and monthly payments.  Use formulas so I can change the loan amount, rate, and term.  Include columns for payment number, payment, principal, interest, and remaining balance.  Add a summary section showing total interest paid.''}
\end{shadedbox}
\vspace{0.2cm}
\begin{itemize}\small
  \item Claude builds the table with \alert{live Excel formulas}, not hardcoded values
  \item Change the rate or loan amount and the entire table recalculates
  \item Ask follow-ups: \textit{``Add a chart of principal vs.\ interest over time''} or \textit{``What if I make an extra \$200/month payment?''}
\end{itemize}
\end{frame}

\begin{frame}{AI Spreadsheet Tools Compared}
\begin{itemize}
  \item \textbf{Claude for Excel} (\$20/mo): formula-preserving edits, no OneDrive required --- best for financial modeling
  \item \textbf{Microsoft Copilot} (\$30/mo): Agent Mode and \texttt{=COPILOT()} function, Python in Excel --- best for enterprise ecosystems
  \item \textbf{Google Sheets + Gemini} (\$20/mo): \texttt{=AI()} function, web-only --- best for cloud collaboration
\end{itemize}
\vspace{0.3cm}
All three can produce the same kinds of financial tools --- DCF calculators, amortization tables, scenario analyzers.  Choose based on your ecosystem.
\end{frame}

% ========================================
% SECTION 4: AI AGENTS
% ========================================
\section{AI Agents}

\begin{frame}{Chatbot}
\begin{center}
\begin{tikzpicture}[
  node distance=1.5cm and 3cm,
  every node/.style={font=\small},
  sbox/.style={draw=titlegray, rounded corners, minimum width=2.5cm, minimum height=1.2cm, fill=excelinput, text=titlegray, font=\normalsize\bfseries, align=center},
  fwd/.style={-{Stealth[length=3mm]}, thick, draw=accentblue},
  bwd/.style={-{Stealth[length=3mm]}, thick, draw=accentblue}
]
\node[sbox] (user) {User};
\node[sbox, right=3cm of user, fill=accentblue!15] (chatbot) {Chatbot};
\node[sbox, right=3cm of chatbot, fill=alertorange!15] (llm) {LLM};

\draw[fwd] ([yshift=2pt]user.east) -- ([yshift=2pt]chatbot.west);
\draw[bwd] ([yshift=-2pt]chatbot.west) -- ([yshift=-2pt]user.east);
\draw[fwd] ([yshift=2pt]chatbot.east) -- ([yshift=2pt]llm.west);
\draw[bwd] ([yshift=-2pt]llm.west) -- ([yshift=-2pt]chatbot.east);
\end{tikzpicture}
\end{center}
\vspace{0.3cm}
The chatbot sends three things to the LLM with each request:
\begin{enumerate}
  \item \textbf{User prompt} --- what you just typed
  \item \textbf{System prompt} --- hidden instructions that define the LLM's behavior, e.g., \textit{``You are a helpful assistant \ldots''}
  \item \textbf{Conversation history} --- all prior messages in the session, so the LLM has context
\end{enumerate}
\vspace{0.3cm}
Anthropic publishes the system prompt for Claude: \href{https://docs.anthropic.com/en/docs/about-claude/soul}{docs.anthropic.com/en/docs/about-claude/soul}
\end{frame}

\begin{frame}{Agent}
\begin{center}
\begin{tikzpicture}[
  node distance=1.5cm and 3cm,
  every node/.style={font=\small},
  sbox/.style={draw=titlegray, rounded corners, minimum width=2.5cm, minimum height=1.2cm, fill=excelinput, text=titlegray, font=\normalsize\bfseries, align=center},
  fwd/.style={-{Stealth[length=3mm]}, thick, draw=accentblue},
  bwd/.style={-{Stealth[length=3mm]}, thick, draw=accentblue}
]
\node[sbox] (user) {User};
\node[sbox, right=3cm of user, fill=accentblue!15] (agent) {Agent};
\node[sbox, right=3cm of agent, fill=alertorange!15] (llm) {LLM};
\node[sbox, below=1.2cm of agent, fill=excelinput] (tool) {Tool};

\draw[fwd] ([yshift=2pt]user.east) -- ([yshift=2pt]agent.west);
\draw[bwd] ([yshift=-2pt]agent.west) -- ([yshift=-2pt]user.east);
\draw[fwd] ([yshift=2pt]agent.east) -- ([yshift=2pt]llm.west);
\draw[bwd] ([yshift=-2pt]llm.west) -- ([yshift=-2pt]agent.east);
\draw[fwd] ([xshift=2pt]agent.south) -- ([xshift=2pt]tool.north);
\draw[bwd] ([xshift=-2pt]tool.north) -- ([xshift=-2pt]agent.south);
\end{tikzpicture}
\end{center}
\vspace{0.2cm}
The agent programmatically controls communication between the LLM and the tool.  The system prompt includes \textit{``You have a tool \ldots''} so the LLM knows what it can call.  Multiple rounds of communication may occur between the LLM, agent, and tool before the agent returns output to the user.
\end{frame}

\begin{frame}{Examples of Agent Tools}
\begin{itemize}
  \item Python environment
  \item Database server
  \item Web browser / web search
  \item File system (read, write, edit)
  \item API calls (REST endpoints, data providers)
\end{itemize}
\vspace{0.5cm}
An agent may be connected to \alert{multiple tools}.  For example, get data from a database server and send it to a Python environment to return analysis.
\end{frame}

\begin{frame}{Example: Rice Data Portal}
The \href{https://data-portal.rice-business.org}{Rice Data Portal} is an agent.
\vspace{0.3cm}
\begin{itemize}
  \item \textbf{LLM:} ChatGPT --- interprets your natural-language question and writes a SQL query
  \item \textbf{Tool:} A database server in the cloud holding stock-market and financial data
  \item \textbf{Agent:} An app running elsewhere in the cloud that coordinates communication between ChatGPT and the database
\end{itemize}
\vspace{0.3cm}
You type a question (\textit{``Show me AAPL closing prices for 2024''}).  The agent sends it to ChatGPT, which returns a SQL query.  The agent runs the query against the database, sends the results back to ChatGPT for formatting, and displays the answer.  If the query fails, ChatGPT rewrites it and tries again --- multiple rounds, no human intervention.
\end{frame}

\end{document}
