\documentclass[aspectratio=169]{beamer}
\usetheme{metropolis}
\usepackage{appendixnumberbeamer}
\usepackage{booktabs, hyperref}

\input{mgmt675-style}

\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows.meta, positioning, calc}

% Title info
\subtitle{MGMT 675: Generative AI for Finance}
\title{Module 3b: Company Valuation}
\author{Kerry Back}
\date{}

\begin{document}

\maketitle

\begin{frame}{Learning Objectives}
\begin{enumerate}
  \item Build two-stage DCF models driven by a small set of assumptions
  \item Perform sensitivity analysis and Monte Carlo simulation to quantify valuation uncertainty
  \item Use AI to build, extend, and stress-test valuation models
\end{enumerate}

\vspace{0.5cm}
These are core tools of corporate finance.  AI handles the mathematics and code.  \alert{You focus on the assumptions and the interpretation.}
\end{frame}

% ============================================================
% COMPANY VALUATION: DCF
% ============================================================
\section{Company Valuation: DCF}

\begin{frame}{The Big Picture}
Discounted cash flow (DCF) valuation estimates the value of a company's operations by forecasting future free cash flows and discounting them to the present.

\vspace{0.3cm}
\begin{itemize}
  \item A small set of \textbf{assumptions} drives the entire model
  \item AI builds the pro forma financial statements and computes free cash flow
  \item You focus on \alert{choosing the right assumptions} and \alert{interpreting the output}
\end{itemize}
\end{frame}

\begin{frame}{The Assumption Set}
A pro forma model is driven by a small number of assumptions.  Everything else is computed from these.
\vspace{0.2cm}
\begin{center}
\small
\begin{tabular}{@{} l l @{}}
\toprule
\textbf{Assumption} & \textbf{Driver} \\
\midrule
Sales growth rate & \% per year \\
COGS & \% of sales \\
SG\&A & Base amount + \% of sales \\
Net working capital (NWC) & \% of sales \\
PP\&E & \% of sales \\
Net other operating assets & \% of sales \\
Depreciation & \% of PP\&E \\
Tax rate & \% of pre-tax income \\
\bottomrule
\end{tabular}
\end{center}
\vspace{0.2cm}
Capital expenditures are not assumed directly.  \alert{Cap ex = target PP\&E $-$ prior PP\&E + depreciation}.
\end{frame}

\begin{frame}{Pro Forma Income Statement}
\begin{center}
\small
\begin{tabular}{@{} l l @{}}
\toprule
\textbf{Line Item} & \textbf{How It's Computed} \\
\midrule
Sales & Prior sales $\times$ (1 + growth rate) \\
COGS & COGS\% $\times$ sales \\
\textbf{Gross profit} & Sales $-$ COGS \\
SG\&A & Base + SG\&A\% $\times$ sales \\
Depreciation & Depr\% $\times$ PP\&E \\
\textbf{EBIT} & Gross profit $-$ SG\&A $-$ depreciation \\
Taxes & Tax rate $\times$ EBIT \\
\textbf{NOPAT} & EBIT $-$ taxes \\
\bottomrule
\end{tabular}
\end{center}
\vspace{0.3cm}
We use \textbf{NOPAT} (net operating profit after tax) rather than net income because DCF values the \textit{operations} of the firm, independent of how they are financed.
\end{frame}

\begin{frame}{Pro Forma Balance Sheet and Cap Ex}
\begin{columns}[T]
\begin{column}{0.48\textwidth}
\begin{shadedbox}[title=\textbf{Balance Sheet Items}]
\begin{itemize}\small
  \item \textbf{NWC} = NWC\% $\times$ sales
  \item \textbf{PP\&E} = PP\&E\% $\times$ sales
  \item \textbf{Net other OA} = Net other OA\% $\times$ sales
  \item Change in each = current $-$ prior year
\end{itemize}
\end{shadedbox}
\end{column}
\begin{column}{0.48\textwidth}
\begin{shadedbox}[title=\textbf{Capital Expenditures}]
\begin{itemize}\small
  \item Target PP\&E$_t$ = PP\&E\% $\times$ sales$_t$
  \item Depreciation$_t$ = depr\% $\times$ PP\&E$_{t-1}$
  \item \alert{Cap ex$_t$ = PP\&E$_t$ $-$ PP\&E$_{t-1}$ + depr$_t$}
\end{itemize}
\end{shadedbox}
\end{column}
\end{columns}
\vspace{0.3cm}
\textbf{Free Cash Flow to the Firm}
\begin{center}
\textbf{FCF = NOPAT + Depreciation $-$ Cap Ex $-$ $\Delta$NWC $-$ $\Delta$Net Other OA}
\end{center}
\end{frame}

% ============================================================
% TWO-STAGE DCF
% ============================================================
\section{Two-Stage DCF}

\begin{frame}{Two-Stage DCF: Overview}
\begin{center}
\begin{tikzpicture}[
  node distance=0.8cm and 2cm,
  every node/.style={font=\small},
  sbox/.style={draw=titlegray, rounded corners, minimum width=3.5cm, minimum height=0.9cm, fill=excelinput, text=titlegray, font=\small\bfseries, align=center},
  sarrow/.style={-{Stealth[length=3mm]}, thick, draw=accentblue}
]
\node[sbox] (s1) {Stage 1\\Explicit Forecast\\(e.g., 5--10 years)};
\node[sbox, right=2cm of s1] (s2) {Stage 2\\Terminal Value\\(perpetuity)};
\node[sbox, below=1.2cm of $(s1)!0.5!(s2)$] (ev) {Enterprise Value\\= PV(Stage 1) + PV(Stage 2)};

\draw[sarrow] (s1) -- (ev);
\draw[sarrow] (s2) -- (ev);
\end{tikzpicture}
\end{center}
\vspace{0.2cm}
\begin{itemize}
  \item \textbf{Stage 1}: Project FCF each year from pro forma assumptions
  \item \textbf{Stage 2}: Terminal value = FCF$_{T+1}$ / (WACC $-$ $g$), where $g$ is the long-run growth rate
  \item \textbf{Equity value} = Enterprise value $-$ net debt (discount at the WACC)
\end{itemize}
\end{frame}

\begin{frame}{Terminal Value}
The terminal value typically accounts for \alert{60--80\%} of total enterprise value.  Getting it right matters more than the explicit forecast.
\vspace{0.3cm}
\begin{columns}[T]
\begin{column}{0.48\textwidth}
\begin{shadedbox}[title=\textbf{Growing Perpetuity}]
\begin{itemize}\small
  \item TV = FCF$_{T+1}$ / (WACC $-$ $g$)
  \item FCF$_{T+1}$ = year $T$ FCF $\times$ (1 + $g$)
  \item $g$ = long-run nominal growth (typically 2--3\%)
  \item Requires $g <$ WACC
\end{itemize}
\end{shadedbox}
\end{column}
\begin{column}{0.48\textwidth}
\begin{shadedbox}[title=\textbf{Exit Multiple}]
\begin{itemize}\small
  \item TV = EBITDA$_T$ $\times$ exit multiple
  \item Multiple from comparable firms
  \item Common in practice (M\&A, PE)
  \item Cross-check against perpetuity method
\end{itemize}
\end{shadedbox}
\end{column}
\end{columns}
\end{frame}

% ============================================================
% SENSITIVITY ANALYSIS
% ============================================================
\section{Sensitivity Analysis}

\begin{frame}{Sensitivity Tables}
A sensitivity table shows how the output (e.g., equity value per share) changes as you vary one or two key inputs.
\vspace{0.3cm}
\begin{columns}[T]
\begin{column}{0.48\textwidth}
\begin{shadedbox}[title=\textbf{One-Way Table}]
\begin{itemize}\small
  \item Vary a single input (e.g., WACC from 8\% to 12\%)
  \item Hold everything else constant
  \item Shows which inputs matter most
\end{itemize}
\end{shadedbox}
\end{column}
\begin{column}{0.48\textwidth}
\begin{shadedbox}[title=\textbf{Two-Way Table}]
\begin{itemize}\small
  \item Vary two inputs simultaneously
  \item Classic: WACC vs.\ terminal growth rate
  \item Also useful for sales growth vs.\ COGS margin; reveals interaction effects
\end{itemize}
\end{shadedbox}
\end{column}
\end{columns}
\vspace{0.3cm}
Ask Claude to produce sensitivity tables in Excel (with formulas) or as formatted output.  In Excel, the Data Table feature automates two-way tables.
\end{frame}

% ============================================================
% MONTE CARLO
% ============================================================
\section{Monte Carlo Simulation}

\begin{frame}{Why Simulate?}
\vfill
\begin{itemize}
  \item Sensitivity tables vary one or two inputs; in reality, \alert{all assumptions are uncertain simultaneously}
  \item Monte Carlo simulation draws random values for each assumption, computes FCF and enterprise value, and repeats thousands of times
  \item The result: a \textbf{distribution} of enterprise values, not a single point estimate
\end{itemize}
\end{frame}

\begin{frame}{Simulation Setup}
Assign a probability distribution to each uncertain assumption, then sample and compute.
\vspace{0.2cm}
\begin{center}
\small
\begin{tabular}{@{} l l l @{}}
\toprule
\textbf{Assumption} & \textbf{Distribution} & \textbf{Example} \\
\midrule
Sales growth & Normal & $\mu = 8\%$, $\sigma = 3\%$ \\
COGS \% & Normal & $\mu = 60\%$, $\sigma = 2\%$ \\
Terminal growth & Uniform & 1.5\% to 3.5\% \\
WACC & Normal & $\mu = 10\%$, $\sigma = 1\%$ \\
\bottomrule
\end{tabular}
\end{center}
\vspace{0.1cm}
\begin{itemize}
  \item Run 10{,}000 simulations $\rightarrow$ 10{,}000 enterprise values
  \item Report mean, median, 10th/90th percentiles; plot a histogram
  \item Identify which assumptions drive the most variance (tornado chart)
\end{itemize}
\end{frame}

% ============================================================
% BUILDING WITH AI
% ============================================================
\section{Building These Analyses with AI}

\begin{frame}{Multiple Ways to Build with AI}
\begin{itemize}
  \item \textbf{Chat}: ``Build a two-stage DCF model for a company with these assumptions \ldots\ Generate an Excel file with pro formas, FCF, and a sensitivity table.''
  \item \textbf{Cowork}: Point Claude at a folder with data files. ``Read the financial data, build pro formas, run a DCF, and save the results to Excel.''
  \item \textbf{Code}: ``Build a DCF model in Python. Fetch Apple's financials from FMP, estimate assumptions from historical data, and run a Monte Carlo simulation.''
  \item \textbf{Excel add-in}: Open a blank workbook. ``Build a two-stage DCF with pro forma statements, a sensitivity table for WACC vs.\ terminal growth, and a tornado chart.''
\end{itemize}
\vspace{0.2cm}
\alert{Key difference:} Chat and Cowork generate spreadsheets via Python (no internet).  Code mode can \alert{fetch live data} from APIs before building the model.
\end{frame}

\begin{frame}{}
\vspace{2cm}
\begin{center}
{\Large\bfseries The same analysis that takes hours with traditional tools takes minutes with AI.}

\vspace{1cm}
Describing the problem clearly---choosing the right assumptions, interpreting the output---is the hard part.  The computation is delegated entirely.
\end{center}
\end{frame}

% ============================================================
% EXERCISES
% ============================================================
\section{Exercises}

\begin{frame}{Exercise 3: DCF Model}
Build a two-stage DCF for a hypothetical company:
\vspace{0.2cm}
\begin{itemize}\small
  \item Current sales: \$500M; sales growth (yrs 1--5): 10\%; terminal growth: 3\%
  \item COGS: 58\% of sales; SG\&A: \$20M + 12\% of sales
  \item NWC: 15\% of sales; PP\&E: 40\% of sales; net other OA: 5\% of sales
  \item Depreciation: 10\% of PP\&E; tax rate: 25\%
  \item WACC: 9\%; net debt: \$200M; shares outstanding: 50M
\end{itemize}
\vspace{0.2cm}
Produce an Excel workbook with pro forma income statement, balance sheet items, FCF calculation, enterprise value, per-share value, and a two-way sensitivity table (WACC vs.\ terminal growth rate).
\end{frame}

\begin{frame}{Exercise 4: DCF with Live Data}
\begin{itemize}
  \item In Claude Code, fetch a real company's financials from the Rice Data Portal (SF1 table) and estimate assumptions from trailing averages.
  \item Build a 5-year pro forma, run a two-stage DCF, and compare to the current market cap.
  \item Submit the code output + a short memo explaining the comparison.
\end{itemize}

\vspace{0.3cm}
\textbf{Note:} This exercise requires Claude Code (internet access for the database query).
\end{frame}

\begin{frame}{Exercise 5: Monte Carlo Simulation}
\begin{itemize}
  \item Take your DCF model and assign distributions to sales growth, COGS \%, terminal growth, and WACC.
  \item Run 10{,}000 simulations; report mean, median, 10th/90th percentiles.
  \item Plot a histogram + tornado chart; discuss which assumption drives the most variation.
\end{itemize}
\end{frame}

\end{document}
